---
title: "Regression Models - Week 4 Course Project"
author: "Sanjay Lonkar"
date: "30 Jun 2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Analyze data set "mtcars" of collection of cars and answer following two questions:


1. Is an automatic or manual transmission better for MPG


2. Quantify the MPG difference between automatic and manual transmissions


## Result

##### Manual transmission is better for MPG than Autaomatic transmission. Also, manual transmission has 1.8 greater MPG than automatic transmission.


### Load data set and perform exploratory data analysis

See Appendix 1 for brief exploratory data analysis.

```{r echo=TRUE}
data (mtcars)
```

### Visual Analysis


###### Verify impact on MPG by generating a box plot of Transmission Type against MPG. See Appendix 2 for generated box plot. It shows MPG is less for Aumotatic transmission type cars.


### Quantified Regression Analysis


```{r echo=TRUE}
aggregate(mpg~am, data = mtcars, mean)
mtcars$cyl  <- factor(mtcars$cyl)
mtcars$vs   <- factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
mtcars$am   <- factor(mtcars$am,labels=c("Automatic","Manual"))
tManual <- mtcars [mtcars$am == "Manual", ]
tAutomatic <- mtcars [mtcars$am == "Automatic", ]
t.test(tAutomatic$mpg, tManual$mpg)
```

It shows that Autmoatic cars have lower 7.25 lower MPG (24.39231 - 17.14737) than Manual cars. Corresponding P-Value is 0.001374 which indicated substantial difference. 

```{r echo=TRUE}
regressionModel <- lm (mpg ~ am, data = mtcars)
summary (regressionModel)
```

Manual car average MPG is 7.2 higher than that of Autmoatic cars which has MPG of 17.1 MPG. Constructing new multivariate regression model as above shows only 36% of the variance (R squared).

```{r echo=TRUE}
multivariateModel <- lm (mpg ~ am + cyl + disp + hp + wt, data = mtcars)
anova (regressionModel, multivariateModel)
```
```{r echo=TRUE, results='hide'}
summary(multivariateModel) # See Appendix 3 for result
```
We have lesser P-Value of 8.637e-08 indicating better model. Pair plot in Appendix 4 shows correlations between difference data points and MPG.

##### Conclusion, new model gives us 86.64% variance. Taking into account disp, hp, cyl and wt affect on correlation between mpg and am, we can say that difference between Manual and Automatic cars MPG is 1.81 MPG.


### Appendix 1

Brief Exploratory Analysis

```{r echo=TRUE}
head (mtcars)
str (mtcars) 
```


### Appendix 2


Box plot visually showing Manual cars give higher MPG than Automatic cars
```{r echo=TRUE}
boxplot(mpg ~ am, data = mtcars, col = (c("red","blue")), ylab = "Miles Per Gallon", 
        xlab = "Transmission Type")
```


### Appendix 3


Summary of multivariate model that gave minimum P-Value
```{r echo=TRUE}
summary(multivariateModel)
```

### Appendix 4


Pair plot indicating correlations between difference data points and MPG

```{r echo=TRUE}
pairs(mpg ~ ., data = mtcars)
```